Reproducible Research_Peer Assessment 1
===============================================
## Loading and preprocessing the data
```{r library, echo = FALSE}
library(ggplot2)
library(plyr)
```
loading data
```{r loaddata, echo = TRUE}
data <- read.table("activity.csv", sep = ",", header = TRUE, na.strings = "NA", colClasses = c("integer", "Date", "factor"))
```

preprocessing data
```{r preprocessing, echo=TRUE}
pdata <- na.omit(data)
```


## The mean total number of steps taken per day

Plot of the total number of steps taken per day
```{r totalsteps, echo=TRUE}
stepsperday <- tapply(pdata$steps, pdata$date, sum)
ggplot(pdata, aes(date, steps)) + geom_histogram(stat = "identity", binwidth = 0.5) + ggtitle("Total Number of Steps Taken Per Day") + xlab("Date") + ylab("Total Number of Steps")
```

Calculate the Mean and Median value
```{r MeannMedian, echo=TRUE}
mean <- mean(stepsperday)
median <- median(stepsperday)
```
####The mean of steps taken per day is: `r mean`
####The median of steps taken per day is: `r median`


## The average daily activity pattern

Plot of the average daily activity pattern
```{r avgactivity, echo=TRUE}
avg <- aggregate(pdata$steps, list(interval = as.numeric(as.character(pdata$interval))), mean)
names(avg)[2] <- "avgsteps"

ggplot(avg, aes(interval, avgsteps)) + geom_line() + ggtitle("Average Daily Activity Pattern") + xlab("5-minute intervals") + ylab("Number of steps")
```

Calculate the maximum avgsteps, and report the interval contains maxinum steps
```{r max, echo=TRUE}
maxsteps <- avg[avg$avgsteps == max(avg$avgsteps),]
maxinterval <- maxsteps[1]
```
####The interval contains maxinum mumber of steps is: `r maxinterval`


## Imputing missing values

Calculate the total number of missing values
```{r missingvalues, echo=TRUE}
missingvalue <- sum(!complete.cases(data))
```
####The total number of missing values is: `r missingvalue`

Imputing missing values by mean value at the same interval across days.
```{r imputedata, echo=TRUE}
impdata <- data
for (i in 1:nrow(impdata)) {
        if (is.na(impdata$steps[i])) {
                impdata$steps[i] <- avg[which(impdata$interval[i] == avg$interval), ]$avgsteps
        }
}
newmissingvalue <- sum(!complete.cases(impdata))
```
####The total number of missing values after imputing is: `r newmissingvalue`

Plot of total number of steps taken eacy day
```{r impdataplot, echo=TRUE}
ggplot(impdata, aes(date, steps)) + geom_histogram(stat = "identity", binwidth = 0.5) + ggtitle("Total Number of Steps Taken Per Day (Impute Data)") + xlab("Date") + ylab("Total Number of Steps")
```

Calculate the Mean and Median value
```{r newmeannmedian, echo=TRUE}
impsteps <- tapply(impdata$steps, impdata$date, sum)
newmean <- mean(impsteps)
newmedian <- median(impsteps)
```
####The mean of steps taken per day is:`r newmean`.
mean of imputed data is the same as that of original data.

####The median of steps taken per day is:`r newmedian`.
median of imputed data is greater than that of original data.



## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r twolevels, echo=TRUE}
days <- factor(format(impdata$date, "%A"))
levels(days) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), weekend = c("Saturday", "Sunday"))
table(days)
```

Panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r panelplot, echo=TRUE}
newavg <- aggregate(impdata$steps, list(interval = as.numeric(as.character(impdata$interval)), weekdays = days), mean)
names(newavg)[3] <- "meanSteps"
library(lattice)
xyplot(newavg$meanSteps ~ newavg$interval | newavg$weekdays, layout = c(1, 2), type = "l", xlab = "Interval", ylab = "Number of Steps")
```

